## Question 1

This question contains some maths equations. If you're unsure about summations either ask your presenter or skip this question. We're not bothered about your maths theory skills, just your R skills!

```{r results="hide", echo=TRUE}
european_countries = c("England", "Germany", "France", "Switzerland", "Spain", "Italy", "Boznia and Herzegovina", "Croatia", "Montenegro", "Sweden")
for(i in european_countries) {
    print(paste("The country" , i, "consists of", nchar(i), "characters."))
}
``` 

The `for` loop above gives us information on the number of letters in a selection of ten European countries.

1. Try and run the `for` loop yourself.

2. Change the above loop, such that we only find out the number of letters in the word Montenegro. Hint: use the square bracket notation, `[]`.

    ```{r}
    for(i in european_countries[9]) {
      print(paste("The country" , i, "consists of", nchar(i),
                  "characters."))
    }
    ```
    
3. Rewrite the for loop, such that we find out the number of letters in the first 5 countries.

    ```{r}
    for(i in european_countries[1:5]) {
      print(paste("The country" , i, "consists of", nchar(i),
                  "characters."))
    }
    ```

4. Harder: Try and change the for loop so that the sentences are ordered alphabetically by country.

    ```{r}
    for(i in sort(european_countries)) {
      print(paste("The country" , i, "consists of", nchar(i),
                  "characters."))
    }
    ```

## Question 2

In the notes, we observed that it was straight forward to
loop through a data set and select the maximum values:

```{r echo=TRUE}
dd = data.frame(x = rnorm(10), y = rnorm(10), z = rnorm(10))

max_cols = numeric(ncol(dd))
for (i in seq_along(dd)) {
  max_cols[i] = max(dd[, i])
}
max_cols
```

 * Alter the above the code to calculate the `mean()` instead of the maximum value
 * Now, calculate the standard deviation (via `sd()`) as well as the mean. 
 You should only have a single loop!
    ```{r}
    means = numeric(ncol(dd))
    sds = numeric(ncol(dd))
    for (i in seq_along(dd)) {
      means[i] = mean(dd[, i])
      sds[i] = sd(dd[, i])
    }
    ```
    
## Solutions

Solutions are contained within this package:
```{r eval=FALSE, echo=TRUE}
vignette("solutions1", package = "jrProgramming")
```


