
```{r setup, include=FALSE, cache=FALSE}
library(knitr)
opts_knit$set(out.format = "latex")
knit_theme$set(knit_theme$get("greyscale0"))

options(replace.assign=FALSE,width=50)

opts_chunk$set(fig.path='figure/graphics-', 
               cache.path='cache/graphics-', 
               fig.align='center', 
               dev='pdf', fig.width=5, fig.height=5, 
               fig.show='hold', cache=FALSE, par=TRUE)
knit_hooks$set(crop=hook_pdfcrop)

knit_hooks$set(par=function(before, options, envir){
  if (before && options$fig.show!='none') {
    #par(mar=c(3,3,2,1),cex.lab=.95,cex.axis=.9, mgp=c(2,.7,0),tcl=-.01, las=1)
  }}, crop=hook_pdfcrop)
```

### Basic functions

Consider the following simpe function
```{r, results="hide"}
v = 5
Fun1 = function() {
  v = 0
  return(v)
}
Fun1()
v
``` 
1. Why does the final line return 5 and not 0.
    ```{r echo=echo, results=results}
    ## Fun1 uses the local variable v
    ```
1. Delete line 3 in the above piece of code. Now change `Fun1` to allow
`v` to be passed as an argument, i.e. we can write `Fun1(5)`. Call
this function to make sure it works.

    ```{r echo=echo, results=results}
    Fun1 = function(v) {
      return(v)
    }
    Fun1(10)
    ```

1. Now make the argument in `Fun1` have a default value of $0$.

    ```{r echo=echo, results=results}
    Fun1 = function(v=0) {
      return(v)
    }
    ``` 

### Default arguments:

```{r}
Fun2 = function(x=10) {
  return(x)
}

Fun3 = function(x) {
  return(x)
}
``` 
1. Why does 
    ```{r eval=FALSE}
    Fun2()
    ```
work, but this raises an error
    ```{r eval=FALSE, echo=c(1,2*echo)}
    Fun3()
    ## Fun1()} has a default argument
    ```
 
1. Change `Fun2` so that it returns $\sqrt{x}$.

    ```{r echo=echo}
    Fun2 = function(x=10) {
      return(sqrt(x))
    }
    ``` 

### `if` statements.

```{r }
Fun4 = function(x) {
  if(x==5) {
    y = 0
  } else {
    y = 1
  }
  return(y)
}
``` 
1. Change `Fun4` so that it:
    * returns 1 if $x$ is positive;
    * returns -1 if $x$ is negative;
    * returns 0 if $x$ is zero.
```{r echo=echo}
Fun4 = function(x) {
  rtn_value = 0
  if(x > 0) {
    rtn_value = 1 
  } else if (x < 0) {
    rtn_value = -1
  }
  return(rtn_value)
}
```   

### `for` loops.

```{r results="hide"}
total = 0
for(i in 1:5) {
  total = total + i
}
total
``` 

The `for` loop above calculates 
\[
\sum_{i=1}^5 i = 1 + 2 + 3 + 4 + 5
\]

1. What is the final value of `total1 in the above piece of code?
    ```{r echo=echo, results=results}
    total
    ```
1. Change the above loop to calculate the following summations:
    \[
\mbox{(i)} \sum_{i=1}^{20} (i+1)  
\]
    ```{r echo=echo, results=results}
    total = 0
    for(i in 1:20) {
      total = total + (i + 1)
    }
    total
    ```
\[
\mbox{(ii)} \sum_{j=-10}^{15} j 
\]
    ```{r echo=echo, results=results}
    total = 0
    for(j in -10:15) {
      total = total + j
    }
total
```


### More `for` loops:

```{r }
a = 2
total = 0
for(blob in a:5) {
  total = total + blob
}
```

1. In the code above, delete line 1. Now put the above code in a function called `Fun5`,
where `a` is passed as an argument, i.e. we can call
`Fun5(1)`

    ```{r echo=echo, results=results}
    Fun5 = function(a) {
      total = 0
      for(blob in a:5){
        total = total + blob
      }
      return(total)
    }
    Fun5(1)
    ```

1. Alter the code so that the `for` loop goes from `a` to
`b`, rather than `a` to $5$. Allow `b` to be passed as an
argument, i.e. we can call `Fun5(1,5)`.
    ```{r echo=echo, results=results}
    Fun5 = function(a, b) {
      total = 0
      for(blob in a:b){
        total = total + blob
      }
      return(total)
    }
    Fun5(1, 5)
    ```

1. Change `Fun5` so that it has default arguments of
`a = 1` and `b = 10`.

    ```{r echo=echo, results=results}
    Fun5 = function(a=1, b=10) {
      total = 0
      for(blob in a:b) {
        total = total + blob
      }
      return(total)
    }
    Fun5(5)
    ``` 



## Solutions

The solutions can be viewed via

```{r eval=FALSE}
library(jrProgramming)
vignette("solutions1", package = "jrProgramming")
```

