```{r, include = FALSE, cache = FALSE}
treat = NULL
library(knitr)
knit_theme$set(knit_theme$get("greyscale0"))

options(replace.assign = FALSE, width = 50)

opts_chunk$set(fig.path = "figure/graphics-",
               cache.path = "cache/graphics-",
               fig.align = "center",
                fig.width = 4, fig.height = 4,
               fig.show = "hold", cache = FALSE, par = TRUE)
knit_hooks$set(crop = hook_pdfcrop)
```

In this question, we are going to use a `for` statement to loop over a large data set and construct some scatter plots. As a note, we could do a lot of this task using facets in **ggplot2**, but it is good practise for your for loop skills and as you'll see in some places using a for loop is benificial. To generate the data, run the following piece of R code

```{r echo=TRUE}
data(experiment, package = "jrProgramming")
head(experiment)
``` 

The data frame `exper` represents an experiment, where we have ten treatments:
$A, B, \ldots, J$ and measurements at some time points. We want to create a
scatter plot of measurement against time, for each treatment type.

1. First we create a scatter plot of one treatment:

    ```{r F1, echo=TRUE, eval=TRUE, tidy=FALSE, message=FALSE, fig.keep="none"}
    library("dplyr")
    library("ggplot2")
    treat_a = filter(experiment, treat == "A")
    ggplot(treat_a, aes(x = time, y = values)) +
      geom_point()
    ``` 
   
 
1. To generate a scatter-plot for each treatment, we need to iterate over the different
treatment types:

    ```{r echo=FALSE}
    treatment = tail(exper, 1)$treat
    group = filter(experiment, treat == treatment)
    ```
    ```{r results='hide', fig.keep='none', tidy=FALSE, echo = TRUE, eval=FALSE}
    for (treatment in unique(experiment$treat)) {
      group = filter(experiment, treat == treatment)
      g = ggplot(group, aes(x = time, y = values)) +
        geom_point()
      print(g)
      readline("Hit return for next plot")
    }
    ``` 
    * What does `unique(experiment$treat)` give?
        ```{r}
        ## It gives all treatments.
        ```
    * In the `for` loop, what variable is changing? What are it's possible values?
        ```{r}
        ## The treat variable is changing.
        ## It goes through the different treatments.
        ```
    * What does the `readline()` function do?
        
        ```{r}
        ## It halts execution, waits for user input
        ```


## Questions

1. We can change the $x$-axis label using the `xlab()` function:
    ```{r fig.keep='none', tidy=FALSE, echo=TRUE}
    ggplot(group, aes(x = time, y = values)) +
      geom_point() +
      xlab("Time")
    ```
    ```{r, fig.keep="none"}
    ggplot(group, aes(x = time, y = values)) +
      geom_point() +
      xlab("Time") +
      ylab("Measurement")
    ```
    Use the `ylab()` function to alter the $y$-axis label.

1. To add a title to a plot we use the `ggtitle()` function, viz:

    ```{r F2, tidy=FALSE, fig.keep="none", echo = TRUE}
    ggplot(treat_a, aes(x = time, y = values)) +
      geom_point() +
      xlab("Time") +
      ylab("Measurement") +
      ggtitle("Treatment")
    ```

    We can combine strings/characters using the `paste()` function,

    ```{r }
    paste("Treatment", treatment)
    ```

    Rather than have a static title, make the title of each plot display the treatment type.

    ```{r fig.keep='none', tidy=FALSE}
    ggplot(group, aes(x = time, y = values)) +
      geom_point() +
      xlab("Time") +
      ylab("Measurement") +
      ggtitle(paste0("Treament ", treatment))
    ``` 

1. The y-axis range should really be the same in all graphics. Use the `ylim()` function 
to fix the range, like below, but using better y-limits. __Hint__: Work out the range before the `for` loop.

    ```{r tidy=FALSE, fig.keep="none", echo = TRUE}
    ggplot(treat_a, aes(x = time, y = values)) +
      geom_point() +
      ylim(-100, 100)
    ```

    ```{r  fig.keep='none', tidy=FALSE}
    r = range(experiment$values)
    ggplot(group, aes(x = time, y = values)) +
      geom_point() +
      xlab("Time") +
      ylab("Measurement") +
      ggtitle(paste0("Treament ", treatment)) +
      ylim(r[1], r[2])
    ``` 


1. At each iteration, use the `message()` function to print the average measurement level across all time points. Look at the message help page for examples of it's use!

    ```{r results='hide', message=FALSE}
    ##Within the for loop have the line
    message(mean(group$values))
    ```

1. On each graph, highlight any observations with a blue point if they are
larger than the mean + standard deviations or less than the mean - 
  standard deviations. You should be using another `geom_point()`.
  __Hint__: You don't need `if` statements here. 
  Just subset your data frame and pass this new data frame to the `geom_point()` function, i.e.
  
    ```{r echo=TRUE, eval = FALSE}
    geom_point(data = outliers, aes(x = time, y = values))
    ```
  
    ```{r fig.keep='none', message=FALSE, tidy=FALSE}
    r = range(experiment$values)
    g = ggplot(group, aes(x = time, y = values)) +
      geom_point() +
      xlab("Time") +
      ylab("Measurement") +
      ggtitle(paste0("Treament ", treatment)) +
      ylim(r[1], r[2])
    print(g)

    ##Calculate the limits
    values = group$values
    message(mean(values))
    upper_lim = mean(values) + sd(values)
    lower_lim = mean(values) - sd(values)
    
    ##Extract the points
    outliers = filter(group, values > upper_lim | values < lower_lim)
    g = g +
      geom_point(data = outliers, aes(x = time, y = values),
                 colour = "red")
    print(g)
    ```
  

1. Suppose we wanted to save individual graphs in a pdf file. We can use the ggsave function like so
    ```{r, echo = TRUE, eval = FALSE}
    ggsave(filename = "treatment.pdf", plot = g)
    ```
To get unique file names, use the `paste0()` function.

1. Put your code, i.e. the `for` loop and plotting commands, in a function which takes the data frame as an argument.


## Solutions

Solutions are contained within this package:
```{r eval=FALSE, echo=TRUE}
vignette("solutions5", package = "jrProgramming")
```


```{r tidy=FALSE}
## FULL SOLUTION
viewgraphs = function(data) {
  for (treatment in unique(data$treat)) {
    group = filter(data, treat == treatment)
    g = ggplot(group, aes(x = time, y = values)) +
      geom_point() +
      xlab("Time") +
      ylab("Measurement") +
      ggtitle(paste0("Treament ", treatment)) +
      ylim(r[1], r[2])

    ## Calculate the limits
    values = group$values
    message(mean(values))
    upper_lim = mean(values) + sd(values)
    lower_lim = mean(values) - sd(values)

    ## Extract the points
    outliers = filter(group, values > upper_lim | values < lower_lim)

    g = g +
      geom_point(data = outliers, aes(x = time, y = values), colour = "red")
    print(g)
    ggsave(filename = paste0("treatment", treatment, ".pdf"),
           plot = g)
    readline("Hit return for next plot")
  }
}
```
