## Question 1

Re-order the lines of code below to create a for loop!

1. message(day)

2. days = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

3. } 

4. for(day in days) {

```{r, eval=FALSE}
days = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
for(day in days) {
  message(day)
}
```

Try changing "day" to "i" wherever "day" is seen. Does it change the output? Explain why

```{r}
# The output will remain the same as the iterator, which acts as a placeholder for the iterator values in the vector, days, is being kept consistent throughout the body of the for loop.
```

## Question 2

In the notes, we observed that it was straight forward to
loop through a data set and select the maximum values:

```{r echo=TRUE}
dd = data.frame(x = rnorm(10), y = rnorm(10), z = rnorm(10))

max_cols = numeric(ncol(dd))
for (i in seq_along(dd)) {
  max_cols[i] = max(dd[, i])
}
max_cols
```

 * Alter the above the code to calculate the `mean()` instead of the maximum value
 * Now, calculate the standard deviation (via `sd()`) as well as the mean. 
 You should only have a single loop!
    ```{r}
    means = numeric(ncol(dd))
    sds = numeric(ncol(dd))
    for (i in seq_along(dd)) {
      means[i] = mean(dd[, i])
      sds[i] = sd(dd[, i])
    }
    ```
    
## Solutions

Solutions are contained within this package:
```{r eval=FALSE, echo=TRUE}
vignette("solutions1", package = "jrProgramming")
```


