## Question 1

Consider the following simple function
```{r echo=TRUE}
v = 5
fun1 = function() {
  v = 0
  return(v)
}
fun1()
``` 

1. Why does the final line return 0 and not 5.
    ```{r }
    ## fun1 uses the local variable v
    ```
2. Delete line 3 in the above piece of code. Now change `fun1()` to allow
`v` to be passed as an argument, i.e. we can write `fun1(5)`. Call
this function to make sure it works.

    ```{r}
    fun1 = function(v) {
      return(v)
    }
    fun1(10)
    ```


## Question 2

```{r echo=TRUE}
fun2 = function(x = 10) {
  return(x)
}

fun3 = function(x) {
  return(x)
}
``` 

1. Why does 
    ```{r echo=TRUE}
    fun2()
    ```
    work, but this raises an error
    ```{r eval=FALSE, echo=TRUE}
    fun3()
    ```

    ```{r}
    ## fun3 expects an argument x, but
    ## we haven't given one and there is
    ## no default.
    ```

2. Change `fun2` so that it returns `x*x`.

    ```{r }
    fun2 = function(x = 10) {
      return(x * x)
    }
    ``` 
    
## Question 3

```{r echo=TRUE}
a = 2
total = 0
for (blob in a:5) {
  total = total + blob
}
```

1. In the code above, delete line 1. Now put the above code in a function called `fun5`,
where `a` is passed as an argument, i.e. we can call
`fun5(1)`

    ```{r }
    fun5 = function(a) {
      total = 0
      for (blob in a:5) {
        total = total + blob
      }
      return(total)
    }
    fun5(1)
    ```

2. Alter the code so that the `for` loop goes from `a` to
`b`, rather than `a` to $5$. Allow `b` to be passed as an
argument, i.e. we can call `fun5(1,5)`.
    ```{r }
    fun5 = function(a, b) {
      total = 0
      for (blob in a:b) {
        total = total + blob
      }
      return(total)
    }
    fun5(1, 5)
    ```

3. Change `fun5` so that it has default arguments of
`a = 1` and `b = 10`.

    ```{r }
    fun5 = function(a=1, b=10) {
      total = 0
      for (blob in a:b) {
        total = total + blob
      }
      return(total)
    }
    fun5(5)
    ``` 
    
## Solutions

Solutions are contained within this package:
```{r eval=FALSE, echo=TRUE}
vignette("solutions2", package = "jrProgramming")
```
