## Question 1

```{r echo=TRUE}
Fun4 = function(x) {
  if(x==5) {
    y = 0
  } else {
    y = 1
  }
  return(y)
}
``` 

Change `Fun4` so that it:

  * returns 1 if $x$ is positive;
  * returns -1 if $x$ is negative;
  * returns 0 if $x$ is zero.

    ```{r }
    Fun4 = function(x) {
      rtn_value = 0
      if(x > 0) {
        rtn_value = 1 
      } else if (x < 0) {
        rtn_value = -1
      }
      return(rtn_value)
    }
    ```   

## Question 2

In practical 1, we did captured both the mean of each column in a sample data frame.
```{r, echo = TRUE}
dd = data.frame(x = rnorm(10),
                y = rnorm(10),
                z = rnorm(10))
means = numeric(ncol(dd))
for(i in seq_along(dd)) {
  means[i] = mean(dd[, i])
}
```

The `message()` command provides us with an easier-to-use, more readable alternative to using `print()`. For instance, 

```{r, echo = TRUE, message = TRUE}
x = 5
message("The value of x is ", x)
```

1. Change the for loop such that it returns a warning message every time the mean goes above 0.3. The message should say "Warning, mean is greater than 0.3".

    ```{r}
    means = numeric(ncol(dd))
    for(i in seq_along(dd)) {
      m = mean(dd[,i])
      if(m > 0.3){
        message("Warning, the mean is greater than 0.3")
      }
      means[i] = mean(dd[, i])
    }
    ```

2. Change the message such that it also tells us the column number of a large mean. For instance, if the mean of columns 2 was > 0.3, it would say "Warning, the mean of column 2 is greater than 0.3"

    ```{r}
    means = numeric(ncol(dd))
    for(i in seq_along(dd)) {
      m = mean(dd[,i])
      if(m > 0.3){
        message("Warning, the mean of column ", i, " is greater than 0.3")
      }
      means[i] = mean(dd[, i])
    }
    ```

3. Convert the code into a function that takes two arguments, the data frame and the limit of the mean. 

    ```{r}
    max_cols_limit = function(data, lim) {
      means = numeric(ncol(dd))
      for (i in seq_along(dd)) {
        m = mean(dd[, i])
        if (m > lim) {
          message("Warning, the mean of column ", i, " is greater than ", lim)
        }
        means[i] = mean(dd[, i])
      }
      return(means)
    }
    max_cols_limit(dd, 0.2)
    ```
    
## Solutions

Solutions are contained within this package:
```{r eval=FALSE, echo=TRUE}
vignette("solutions3", package = "jrProgramming")
```